Exposure Consequences - Statistics and Figures

### 2018-07-25 ManuscriptMaterials.Rmd
updates to reflect extra participants and low performance on blink detection

1. Setup & Task (goal: visualize the experiment)
    + A: Setup (side view)
    + B: Rotated training (top view)
    + C: No-cursor reaches (top view)
    + D: Localization (top view)
2. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
    + A: Exposure reach aftereffects (initial & later)
    + B: Classic & Exposure reach aftereffects (initial only? or all?)
3. Localization
    + A: Exposure, active and passive (goal: small/no difference between active and passive)
    + B: Classic, active and passive (goal: larger difference between active and passive, different generalization?)
    + C: Online localization?

@@

```{r loadsources}
source('shared.R') # functions used everywhere
source('nocursor.R') # functions for no-cursor reach data
source('localization.R') # functions for localization data
source('relateLocalizationNoCursors.R') # functions that correlate the two kinds of data
```

@@

In the manuscript we'll first show that there indeed are motor changes after exposure training (with the no-cursor data) and then compare that with the classic. Second is the localization: a) localization shifts after exposure b) the shifts are (not) different from those after classic c) the effects are (not) different across the workspace d) this pattern is (not) different from that after classic. A third topic might be if localization shifts can predict no-cursor changes, which we'll call the "correlation"  topic. Here we'll also follow this order:

1. Reach aftereffects
2. Localization
3. Correlations

@@

Since in the localization tasks there is missing data at some points (usually at the extremes) we can't use ANOVAs directly on the data. Imputation is going to be very hard because the missing data is systematically missing. Hence, we are using Linear Mixed Effects models as an alternative, but this presents a new problem. Getting the usual F-values and p-values is not so straightforward. Initially we decided to use the $\chi^2$ approach that we get from the `Anova` function from the `car` package. However, a paper from Steven G. Luke (2017, _Evaluating significance in linear mixed-effects models in R_) suggests using Sattertwaite approximation. So all functions running an LME now take an argument that tells them to use the Satterthwaite method, but the default is chi-squared. It turns out that it doesn't matter much, so it's a good sanity check. For consistency across all the analyses, we set the value of that argument is set here:

```{r}
LMEmethod <- 'Satterthwaite'
# options: Satterthwaite, chi-squared
```

@@

@@

Here is a plot, probably Figure 2:

```{r fig2_reachaftereffect, fig.width=7.5, fig.height=2.5}
plotReachAftereffects()
```

@@

First, we show that no-cursor reaches change direction immediately after exposure training (iteration 1):

```{r RAE_exposure,include=TRUE}
exposureNoCursorChange(LMEmethod=LMEmethod)
```

@@

```{r RAE_exp_iterations}
exposureAftereffectsPersistent(LMEmethod=LMEmethod)
```

So there is no effect of iteration here, nor does it interact with target, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations. We now also have enough power to see an effect of target angle: the generalization curve is not flat.

@@

Finally, we want to see if there is a difference between the reach aftereffects observed after exposure training and those after classic training.

```{r RAE_exp_cla}
exposureClassicReachAftereffects(LMEmethod=LMEmethod)
```

I'd have expected there to be some difference between the two training types, perhaps only in interaction with target, but that's not the case apparently.

The absence of an interaction between training and target could be interpreted to mean that the generalization of reach aftereffects is the same after both kinds of training. Here are the confidence intervals for the peaks of both:

```{r RAE_peaks}
getPeakConfidenceInterval('classic', part='all', CIs=c(.95))
getPeakConfidenceInterval('exposure', part='all', CIs=c(.95))
```

The 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target. Something might be going on -- but the two confidence intervals are not different from each other, which is in line with the LME results. We can't conclude much, except that we have no evidence for different generalization curves.

@@

Here is a plot, probably figure 3:

```{r fig3_localization, fig.width=7.5, fig.height=2.5}
plotLocalization(classicOnline=TRUE)
```

We first want to see if there is any overall effect of aligned versus rotated training in the exposure group (i.e.: are there any localization shifts at all?), with movement type and hand angle as fixed effects, and particiant as random effect.

```{r loc_exp}
exposureLocalization(remove15=TRUE, LMEmethod=LMEmethod)
```

```{r loc_exp_movtype}
exposureLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod)
```

```{r loc_ombnibus}
groupLocalization(model='full', remove15=TRUE, LMEmethod=LMEmethod)
```

In the exposure group there is no effect of movement type, but we see one here. This could be because the effect of movement type is much larger in the classic group. This would predict an interaction between group and movement type, but that's not the case (p=.133).

What if we don't care about hand angle:

```{r loc_groups_movementtype}
groupLocalization(model='movementtype', remove15=TRUE, LMEmethod=LMEmethod)
```

Still no effect...

So there is a fairly strong effect of movement type in the classic group. Together with the absence of an effect of movement type in the exposure group, but no interaction between group and movement type if we combine both datasets.

Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it seems that after classic training the generalization curve for delayed localization is much flatter.

2. In the exposure group, we find no difference between localization responses with and without efference copies.
3. In the classic group, the localization responses informed by efference copies, are shifted more than those with only proprioception (as found before).
4. This predicts an interaction between movement type and group in the combined analysis, but we don't see that.

```{r localization_peaks}
getPeakLocConfidenceInterval(group='classic',CIs=c(.95), movementtype='active', LRpart='all')
getPeakLocConfidenceInterval(group='exposure',CIs=c(.95), movementtype='active', LRpart='all')
```

These results mimic those of the reach aftereffects: the median values are different, but they each fall in the others confidence interval. However, here the LME showed that there was an interaction between training type and hand angle, which suggests that the generalization curves are different. If we want to look at this better, we need to get better estimates of these curves in every participant, by having a wider range of points and more data per point, but probably also larger rotations.

@@

```{r scattergrams, fig.width=7.5, fig.height=2.5}
correlations <- correlateNoCursorsLocalization()
```

It looks like for exposure, active localization is the better predictor of reach aftereffects, whereas in classic it is passive localization. But are any of them real? The function also returns the correlation tests and linear regressions for all four datasets.

@@

Only one is significant: active localization predicting reach aftereffects in the exposure group.

@@

Here I do step-wise multiple regression, allowing active and passive localization shift as predictor of reach aftereffects. If I do this on all the no-cursor block data, the models converge on using the active localization for the exposure group, and passive localization for the classic group.

If we look at only the initial no-cursor block, the story changes. It uses both active and passive to predict reach aftereffects for the exposure group, but converges on using only an intercept term for the classic group. This might mean that in te classic group there is no relationship between reach aftereffects and localization, or this might mean that in the exposure group we have sufficient amount of measurements to detect the relationship, as we have twice the number of localization trials. One way to rule this out is to use only the first half of the localization data from the exposure group.

@@

```{r step_exposure}
multipleRegressionLocalization(group='exposure',NCpart='initial')
```

No terms are dropped an the adjusted R squared of the full model is low, but reasonable. _Note that the coefficient for the passive localization is positive._

@@

```{r step_classic}
multipleRegressionLocalization(group='classic',NCpart='initial')
```

@@

```{r step_exposure_firsthalf}
multipleRegressionLocalization(group='exposure',NCpart='initial', LRpart='first')
```

Now the passive localization shifts are thrown out by the algorithm and the active localization shifts don't contribute much either. This means that perhaps previous papers didn't find any relationships between recalibrated proprioception and reach aftereffects because not enough data was collected to overcome the noisiness of the measurements?

@@

```{r step_online}
multipleRegressionLocalization(group='online',NCpart='initial')
```

But no.

@@

```{r create_svgs, cache=FALSE, include=FALSE}
# This chunk creates SVG files but shouldn't show up in the PDF / HTML 
plotReachAftereffects(generateSVG=TRUE)
plotLocalization(classicOnline=TRUE, generateSVG=TRUE)
correlations <- correlateNoCursorsLocalization(generateSVG=TRUE)
```

>>>>

1. Setup & Task (goal: visualize the experiment)
    + A: Setup (side view)
    + B: Rotated training (top view)
    + C: No-cursor reaches (top view)
    + D: Localization (top view)
2. Localization
    + A: Exposure, active and passive (goal: small/no difference between active and passive)
    + B: Classic, active and passive (goal: larger difference between active and passive, different generalization?)
    + C: Online localization?
3. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
    + A: Exposure reach aftereffects (initial & later)
    + B: Classic & Exposure reach aftereffects (initial only? or all?)

@@

```{r loadsources}
source('shared.R') # functions used everywhere
source('localization.R') # functions for localization data
source('nocursor.R') # functions for no-cursor reach data
source('relateLocalizationNoCursors.R') # functions that correlate the two kinds of data
```

@@

In the manuscript we'll first show that there indeed are localization shifts after exposure training (with the no-cursor data) and then compare that with the classic. Second are the no-cursor reaches (and the reach aftereffects). A third topic might be if localization shifts can predict no-cursor changes, which we'll call the "correlation" topic. Here we'll also follow this order:

1. Analyses (some general remarks about analyses)
2. Participants (participant demographics)
3. Localization
4. Reach aftereffects
5. Correlations

@@

Since in the localization tasks there is missing data at some points (usually at the extremes) we can't use ANOVAs directly on the data. Imputation is going to be very hard because the missing data is systematically missing. Hence, we are using Linear Mixed Effects models as an alternative, but this presents a new problem. Getting the usual F-values and p-values is not so straightforward. Initially we decided to use the $\chi^2$ approach that we get from the `Anova` function from the `car` package. However, a paper from Steven G. Luke (2017, _Evaluating significance in linear mixed-effects models in R_) suggests using Satterthwaite approximation. So all functions running an LME now take an argument that tells them to use the Satterthwaite method, but the default is chi-squared. It turns out that it doesn't matter much, so it's a good sanity check. For consistency across all the analyses, we set the value of that argument here:

```{r}
LMEmethod <- 'Satterthwaite'
# options: 'Satterthwaite', 'chi-squared'
```

@@

# Participants

There are no scripts for this part, as the code is simple and can just be embedded.

## Performance

Let's look at the blink detect performance.

```{r}
plotBlinkDetection()
```

Turns out the blink detection is hard and needs practice. But three participants are still performing around chance in the rotated training task, so we want to remove them from the data:

```{r}
blinks <- load.DownloadDataframe(informationURLs['blinkdetect'],'blinkdetect_exposure.csv')
OKparticipants <- blinks$participant[which(blinks$rotated_b == 1 & blinks$performance > 0.65)]
```

We'll also set a variable that is to be passed to all functions to exclude the underperformance:

```{r}
selectPerformance <- TRUE
```

Which can also be set to FALSE to include those three participants in all analyses and plots.

## Demographics

First load the data, and select only the well performing ones:

```{r eval=F}
participants <- load.DownloadDataframe(informationURLs['demographics'],'participants_exposure.csv')
participants <- participants[which(participants$participant %in% OKparticipants),]
```

We want to know some demographics, first mean age, and standard deviation:

```{r eval=F}
c('mean'=mean(participants$age), 'sd'=sd(participants$age))
```

And the distribution of sex:

```{r eval=F}
table(participants$sex)
```

Here is a plot, probably figure 2:

```{r fig3_localization, fig.width=7.5, fig.height=2.5}
plotLocalization(classicOnline=TRUE,selectPerformance=selectPerformance)
```

## Missing data

Here are the counts of participants with valid localization estimates in all 4 tasks for every hand angle. Missing data occurs here, because we don't extrapolate beyond the range of the data.

```{r localization_count_table}
kable(getLocCountTable(selectPerformance=selectPerformance))
```

Since there is a good amount of missing data at the 15 degree location, we exclude that angle from analysis, with a switch in analysis functions `remove15=TRUE`, which can be set to false if we'd want.

We first want to see if there is any overall effect of aligned versus rotated training in the exposure group (i.e.: are there any localization shifts at all?), with movement type and hand angle added as fixed effects as well, and particiant as random effect.

```{r loc_exp}
exposureLocalization(remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
```

```{r loc_exp_movtype}
exposureLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
```

```{r loc_ombnibus}
groupLocalization(model='full', remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
```

In the exposure group there is no effect of movement type, but we see one here. This could be because the effect of movement type is much larger in the classic group. This would predict an interaction between group and movement type, and we do see that (p=.035). (Including the three)

Perhaps hand angle factored into this somehow, so let's see what happens if we remove it:

```{r loc_groups_movementtype}
groupLocalization(model='movementtype', remove15=TRUE, LMEmethod=LMEmethod, selectPerformance=selectPerformance)
```

Still an effect...

So there is a reliable effect of movement type in the classic group, which is in line with the interaction between group and movement type.

Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it does appear that after classic training the generalization curve for delayed localization is much flatter.

2. In the exposure group, we find no difference between localization responses with and without efference copies.
3. In the classic group, the localization responses informed by efference copies, are shifted more than those with only proprioception (as found before).
4. This predicts an interaction between movement type and group in the combined analysis, and we see that.

There was an interaction between group and hand angle on localization shifts above, which suggests different generalization curves in the two groups. This could reflect a difference in the shape (more flat for the classic training group as compared to exposure), but also a difference in the locations of the peaks of the generalization curves.

```{r localization_peaks}
getPeakLocConfidenceInterval(group='classic',
                             CIs=c(.95), 
                             movementtype='active', 
                             LRpart='all',
                             selectPerformance=FALSE)
getPeakLocConfidenceInterval(group='exposure',
                             CIs=c(.95), 
                             movementtype='active', 
                             LRpart='all', 
                             selectPerformance=selectPerformance)
```

The median values are a little over 13 degrees apart, and they _just_ fall outside each others confidence interval. This can explain the interaction between group and hand angle on localization shifts.

@@

Here is a plot, probably Figure 3:

```{r fig2_reachaftereffect, fig.width=7.5, fig.height=2.5}
plotReachAftereffects(selectPerformance=FALSE)
```

@@

First, we look at the no-cursor reaches after exposure training only (see Fig 3a). We show that no-cursor reaches change direction immediately after exposure training (iteration 1):

```{r RAE_exposure,include=TRUE}
exposureNoCursorChange(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
```

@@

```{r RAE_exp_iterations}
exposureAftereffectsPersistent(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
```

So there is no effect of iteration here, nor does it interact with target, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations. We now also see an effect of target angle: the generalization curve is not flat. This seems mostly due to the extra training.

@@

Finally, we want to see if there is a difference between the reach aftereffects observed after exposure training and those after classic training. Here we combine the data across all iterations (see Fig 3b).

```{r RAE_exp_cla}
exposureClassicReachAftereffects(LMEmethod=LMEmethod, selectPerformance=selectPerformance)
```

The figure looked like there might be some difference between the two training types, mainly in where the generalization curves peak. There is no main effect of training type, so the overall magnitude of reach aftereffects is not different. But there is an interaction between training and target (p=.038), so that perhaps the curves peak at different locations.

So we now investigate the peaks the confidence intervals for the peaks of both:

```{r RAE_peaks}
getPeakConfidenceInterval('classic', part='all', CIs=c(.95), selectPerformance=selectPerformance)
getPeakConfidenceInterval('exposure', part='all', CIs=c(.95), selectPerformance=selectPerformance)
```

The 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target. Something might be going on -- but the two confidence intervals are not different from each other, which is not in line with the LME results. We can't conclude much, except that we don't have strong evidence for different generalization curves.

@@

@@

```{r scattergrams, fig.width=7.5, fig.height=2.5}
correlations <- correlateNoCursorsLocalization(selectPerformance=selectPerformance)
```

It looks like for exposure, both types of localization might be able to predict of reach aftereffects, whereas in classic it is only passive localization. But are any of these reliable? The function also returns the correlation tests and linear regressions for all four datasets.

@@

None are significant... some are "trending".

@@

Here I do step-wise multiple regression, allowing active and passive localization shift as predictor of reach aftereffects.

```{r step_exposure}
multipleRegressionLocalization(group='exposure', 
                               NCpart='all', 
                               selectPerformance=selectPerformance)
```

The algorithm decides to retain active localization as predictor, but the regression model isn't very good. The problem could be that we need no-cursor data and localization data that are both recorded immediately after training. This also makes it more comparable to the classic group that had fewer blocks of data.

But if we look at only the initial no-cursor block, the model doesn't change much. It still uses only active localization and the p-value goes down a bit and the R-squared goes up a bit. For the classic group...

converges on using only an intercept term for the classic group. This might mean that in te classic group there is no relationship between reach aftereffects and localization, or this might mean that in the exposure group we have sufficient amount of measurements to detect the relationship, as we have twice the number of localization trials. One way to rule this out is to use only the first half of the localization data from the exposure group.

@@

```{r step_exposure_initial}
multipleRegressionLocalization(group='exposure',
                               NCpart='initial',
                               LRpart='both', 
                               selectPerformance=selectPerformance)
```

@@

```{r step_classic_initial}
multipleRegressionLocalization(group='classic',
                               NCpart='initial',
                               LRpart='both')
```

@@

```{r step_exposure_firsthalf}
multipleRegressionLocalization(group='exposure',
                               NCpart='initial', 
                               LRpart='first', 
                               selectPerformance=selectPerformance)
```

Now we have an intercept model for the exposure group too. Perhaps this means that previous papers didn't find any relationships between recalibrated proprioception and reach aftereffects because not enough data was collected to overcome the noisiness of the measurements?

Maybe including more data for the model for classic might work better:

```{r step_classic_all}
multipleRegressionLocalization(group='classic',
                               NCpart='all',
                               LRpart='both')
```

And now we get the passive localization as in the initial regressions.

@@

```{r step_online}
multipleRegressionLocalization(group='online',
                               NCpart='all')
```

And that might be the case as one term is preserved by the model, but on the other hand, the predictions are pretty bad.

@@

```{r create_svgs, cache=FALSE, include=FALSE}
# This chunk creates SVG files but shouldn't show up in the PDF / HTML 
plotReachAftereffects(generateSVG=TRUE)
plotLocalization(classicOnline=TRUE, generateSVG=TRUE)
correlations <- correlateNoCursorsLocalization(generateSVG=TRUE)
#plotBlinkDetection(generateSVG=TRUE)
```
